QPLIB: A Library of Quadratic Programming Instances

### 2017-02-01 format.tex
add missing [2]
update to qplib format and that pesky 1/2

\newcommand{\vect}[1]{\left(\begin{array}{c}#1\end{array}\right)}
\newcommand{\mat}[2]{\left(\begin{array}{#1}#2\end{array}\right)}
\newcommand{\arr}[2]{\begin{array}{#1}#2\end{array}}
\newcommand{\mininx}[1]{ {\renewcommand{\arraystretch}{0.8}
                     \begin{array}[t]{c}
                     \mbox{minimize} \vspace*{-1mm} \\
                     \mbox{ $\scriptstyle x \in \smallRe^n #1 $ }
                     \end{array} \;} }
\newcommand{\s}[1]{^{\mbox{\protect\tiny #1}}}
\newcommand{\sfrac}[2]{{\scriptstyle \frac{#1}{#2}}}
\newcommand{\half}{\sfrac{1}{2}}
\newcommand{\smallRe}
     {\mbox{\raisebox{-0.8pt}{\footnotesize I\hskip -1.5pt R\hskip -0.5pt}}}
\newcommand{\tim}[1]{\;\; \mbox{#1} \;\;}
\newcommand{\st}{\mbox{subject to}\;\;}
\newcommand{\lbox}[1]{\parbox[t]{105mm}{#1}\vspace*{1mm}}

@@

The QPLIB format is defined in Table~\ref{tab-qplib-format}, with the notation of \S~\ref{sec:QPbasic}.

@@

$n^{Q^i}$ & number of nonzeros (integer) in lower triangle of

@@

& constraint, row and column indices (integers) and value (real) for
 each entry of $Q^{i}$, if $n^{Q^i}$, one quadruple on each line & \\

@@

$n^{b^i}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\
$i$\; $j$\; $b^i_{j}$ & row and column indices (integers) and value (real)
for each nonzero
entry of $b^i$, if $n^{b} > 0$, one triple on each line & \\

@@

in $c_{u,d}$, if $n^{c_{u,d}} > 0$, one pair per line        & \\

@@

 1 for integer variables) & [5] \\

@@

\renewcommand{\descriptionlabel}[1]{\hspace{\labelsep}\texttt{#1}}
\begin{description}
\item [{[1]}]
%\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}

The problem type is represented by a character string as one of the following:
\begin{description}
\item For continuous problems (i.e., all variables are continuous):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [LP]      a linear program,
\item [LPQC]    a linear program with quadratic constraints,
\item [BQP]     a bound-constrained quadratic program,
\item [QP]      a quadratic program, and
\item [QPQC]    a quadratic program with quadratic constraints.
\end{description}
\item For integer problems (i.e., all variables are integer valued):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [ILP]     an integer linear program,
\item [ILPQC]   an integer linear program with quadratic constraints,
\item [IBQP]    an integer bound-constrained quadratic program,
\item [IQP]     an integer quadratic program, and
\item [IQPQC]   an integer quadratic program with quadratic constraints.
\end{description}
\item For mixed-integer problems (i.e., there is a mix of continuous and
integer variables):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [MILP]    a mixed-integer linear program,
\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
\item [MIQP]    a mixed-integer quadratic program, and
\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
\end{description}
\end{description}

\item [{[2]}]
for bound-constrained QPs, these sections are omitted.
\item [{[3]}]
for linear program with quadratic constraints, this section is omitted.
\item [{[4]}]
for problems without quadratic constraints, this section is omitted.
\item [{[5]}]
for purely-continuous or purely-integer problems, this section is omitted.
\end{description}

As a simple example, consider the mixed-integer QP
\[\arr{l}{\min_{x \in \smallRe^3} x_1^2 + x_2^2 + x_3^2 - x_1 x_2 - x_2 x_3
  - 0.2 x_1  -0.4 x_2 - 0.2 x_3 \\
\st 1 \leq x_1 + x_2, \; 1 \leq x_1 + x_3, \; 0 \leq x_1 \leq 1,
\; 0 \leq x_2 \leq 2, \; \mbox{and binary} \; x_3.}
\]
This may then be represent in QPLIB format as follows:

{\small
\begin{verbatim}
! ---------------
! example problem
! ---------------
MIPBAND   # problem name
MIQP      # problem is a mixed-integer quadratic program
Minimize  # minimize the objective function
3         # variables
2         # general linear constraints
5         # nonzeros in lower triangle of Q^0
1 1 2.0   5 lines row & column index & value of nz. in lower triangle Q^0
2 1 -1.0  |
2 2 2.0   |
3 2 -1.0  |
3 3 2.0   |
-0.2      default value for entries in b_0
1         # non default entries in b_0
2 -0.4    1 lines of index & value of non-default values in b_0
0.0       value of q^0
4         # nonzeros in vectors b^i (i =0,...,m)
1 1 1.0   4 lines row & column index & value of nz. in b^i (i =0,...,m)
1 2 1.0   |
2 1 1.0   |
2 3 1.0   |
1.0E+20   infinity
1.0       default value for entries in c_l
0         # non default entries in c_l
1.0E+20   default value for entries in c_u
0         # non default entries in c_u
0.0       default value for entries in l
0         # non default entries in l
1.0       default value for entries in u
1         # non default entries in u
2 2.0     1 line of non-default indices and values in u
0         default variable type is continuous
1         # non default variable types
3 1       variable 3 is integer
1.0       default value for initial values for x
0         # non default entries in x
0.0       default value for initial values for y
0         # non default entries in y
0.0       default value for initial values for z
0         # non default entries in z
0         # non default names for variables
0         # non default names for constraints
\end{verbatim}
}

>>>>

\newcommand{\vect}[1]{\left(\begin{array}{c}#1\end{array}\right)}
\newcommand{\mat}[2]{\left(\begin{array}{#1}#2\end{array}\right)}
\newcommand{\arr}[2]{\begin{array}{#1}#2\end{array}}
\newcommand{\mininx}[1]{ {\renewcommand{\arraystretch}{0.8}
                     \begin{array}[t]{c}
                     \mbox{minimize} \vspace*{-1mm} \\
                     \mbox{ $\scriptstyle x \in \smallRe^n #1 $ }
                     \end{array} \;} }
\newcommand{\s}[1]{^{\mbox{\protect\tiny #1}}}
\newcommand{\smallRe}
     {\mbox{\raisebox{-0.8pt}{\footnotesize I\hskip -1.5pt R\hskip -0.5pt}}}
\newcommand{\tim}[1]{\;\; \mbox{#1} \;\;}
\newcommand{\st}{\mbox{subject to}\;\;}
\newcommand{\lbox}[1]{\parbox[t]{105mm}{#1}\vspace*{1mm}}

@@

The QPLIB format is defined in Table~\ref{tab-qplib-format}, with the notation of \S\ref{sec:QPbasic}.

@@

$n^{Q^i}$ & total number of nonzeros (integer) in lower triangles of

@@

& i, row and column indices (integers) and value (real) for
 each entry of $Q^{i}$ for every $i \in \Mcm$,  if $n^{Q^i}$,
 one quadruple on each line & \\

@@

$n^{b^i}$ & total number of nonzeros  (integer) in $b^i$ for all $i \in \Mcm$ &
 [2] \\
$i$\; $j$\; $b^i_{j}$ & i and index (integers) and value (real)
for each nonzero entry of $b^i$ for every $i \in \Mcm$, if $n^{b} > 0$,
one triple on each line & \\

@@

term in $c_{u,d}$, if $n^{c_{u,d}} > 0$, one pair per line        & [2] \\

@@

 1 for integer variables, 2 for binary variables) & [5] \\

@@

\renewcommand{\descriptionlabel}[1]{\hspace{\labelsep}\texttt{#1}}
\begin{description}
\item [{[1]}]
%\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}
The problem type is represented by a three character string as given
in \S\ref{ssec:taxonomy}
%\begin{description}
%\item For continuous problems (i.e., all variables are continuous):
%\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
%\item [LP]   a problem with linear objective and linear constraints,
%\item [QCP]  a problem with linear objective and linear and quadratic constraints,
%\item [BQP]  a problem with quadratic objective and variable bounds,
%\item [QP]   a problem with quadratic objective and linear constraints, and
%\item [QCQP] a problem with quadratic objective and linear and quadratic constraints.
%\end{description}
%\item For integer problems (i.e., all variables are integer valued):
%\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
%\item [ILP]   an LP plus integer requirement for all variables,
%\item [IQCP]  an QCP plus integer requirement for all variables,
%\item [IBQP]  an BQP plus integer requirement for all variables,
%\item [IQP]   an QP plus integer requirement for all variables, and
%\item [IQCQP] an QCQP plus integer requirement for all variables.
%\end{description}
%\item For mixed-integer problems (i.e., there is a mix of continuous and
%integer variables):
%\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
%\item [MILP]   an LP plus integer requirement for some variables,
%\item [MIQCP]  an QCP plus integer requirement for some variables,
%\item [MIBQP]  an BQP plus integer requirement for some variables,
%\item [MIQP]   an QP plus integer requirement for some variables, and
%\item [MIQCQP] an QCQP plus integer requirement for some variables.
%\end{description}
%\end{description}

\item [{[2]}]
for problems of type \textit{**N} or \textit{**B}, these lines/sections
are omitted.
\item [{[3]}]
for problems of type \textit{L**}, this section is omitted.
\item [{[4]}]
for problems of type \textit{**N}, \textit{**B} or \textit{**L},
this section is omitted.
\item [{[5]}]
for problems of type \textit{*C*}, \textit{*B*} or \textit{*I*},
this section is omitted.
\end{description}

As a simple example, consider the mixed-integer QP
\[\arr{l}{\min_{x \in \smallRe^3} x_1^2 + x_2^2 + x_3^2 - x_1 x_2 - x_2 x_3
  - 0.2 x_1  -0.4 x_2 - 0.2 x_3 \\
\st 1 \leq x_1 + x_2, \; 1 \leq x_1 + x_3, \; 0 \leq x_1 \leq 1,
\; 0 \leq x_2 \leq 2, \; \mbox{and binary} \; x_3,}
\]
for which the Hessian of the objective function is
\[Q^0 = \mat{ccc}{2 & -1 & 0 \\ -1 & 2 & -1 \\ 0 & -1 & 2}.\]
This may then be represent in QPLIB format as follows:

{\small
\begin{verbatim}
! ---------------
! example problem
! ---------------
MIPBAND   # problem name
QML       # problem is a mixed-integer quadratic program
Minimize  # minimize the objective function
3         # variables
2         # general linear constraints
5         # nonzeros in lower triangle of Q^0
1 1 2.0   5 lines row & column index & value of nonzero in lower triangle Q^0
2 1 -1.0  |
2 2 2.0   |
3 2 -1.0  |
3 3 2.0   |
-0.2      default value for entries in b_0
1         # non default entries in b_0
2 -0.4    1 line of index & value of non-default values in b_0
0.0       value of q^0
4         # nonzeros in vectors b^i (i=0,...,m)
1 1 1.0   4 lines constraint, index & value of nonzero in b^i (i=0,...,m)
1 2 1.0   |
2 1 1.0   |
2 3 1.0   |
1.0E+20   infinity
1.0       default value for entries in c_l
0         # non default entries in c_l
1.0E+20   default value for entries in c_u
0         # non default entries in c_u
0.0       default value for entries in l
0         # non default entries in l
1.0       default value for entries in u
1         # non default entries in u
2 2.0     1 line of non-default indices and values in u
0         default variable type is continuous
1         # non default variable types
3 2       variable 3 is binary
1.0       default value for initial values for x
0         # non default entries in x
0.0       default value for initial values for y
0         # non default entries in y
0.0       default value for initial values for z
0         # non default entries in z
0         # non default names for variables
0         # non default names for constraints
\end{verbatim}
}

### 2017-02-01 methods.tex
update to qplib format and that pesky 1/2

In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (\S~\ref{subsec:solver}). For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in \S~\ref{sec:classification}. We remark that many solvers implement more than one algorithm, among which the user can choose at runtime. Moreover, algorithms are typically implemented by different solvers in different ways, so that the same conceptual algorithm can sometimes yield wildly different results or performance measures on the same instances.

>>>>

In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (\S\ref{subsec:solver}). For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in \S\ref{sec:classification}. We remark that many solvers implement more than one algorithm, among which the user can choose at runtime. Moreover, algorithms are typically implemented by different solvers in different ways, so that the same conceptual algorithm can sometimes yield wildly different results or performance measures on the same instances.
